---
title: "Mapping the World"
subtitle: "Using the sf and spData libraries to create simple maps of the Earth"
author: "Isaac Rand"
date: "9 April 2020"
output: 
  html_document:
    theme: yeti
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Setup

### Version

This tutorial was generated using the R version 3.5.3, using a different version may lead to unforeseen issues. You can check your version using output of the code chunk below. 

```{r}
version
```

### Libraries

To follow along and make your own map of a world or countries within it, you will need to load several R libraries. 

If this is your first time using them, you will need to install them first using:

```{r, eval=FALSE}
# if you don't need to install packages skip this part :)

install.packages("sf")
install.packages("spData")
```

Now that the packages are installed, load them now. This will step will need to be repeated every time you begin a new R session

```{r}

library(sf)
library(spData)
```


## Mapping the World

The spData library comes preloaded with an object called `world`. Each row of this object represents a country, complete with its geometry and several pieces of information about it. 

Running `plot(world)` generates, by default, maps of only the first 9 attributes listed for each country (you can change this by running `max.plot = your_desired_number`). 

```{r}
plot(world)
```

Most of the time, we will want to highlight a single variable, not all of them at once. You can select a single variable to plot by appending `["variable_name"]` to the geometry object, as shown below.

```{r}
plot(world["lifeExp"])
```

This method can be used to make simple maps of any of the attributes stored in `world`,including any attributes which you add yourself! Additionally, very similar methods can be used to plot any `sfg` object which you can create yourself or import from a spatial file. You can learn more about creating `sfg` objects in section 2.2.6 of Robin Lovelace's [Geocomputation with R](https://geocompr.robinlovelace.net/spatial-class.html).
